Body size patterns

### 2019-07-12 MS_Analysis_3spp_4rivers_Patterns_from 2spp_bySpp.Rmd
took T and F out of models

mod6 <- bam(observedLength ~ riverFactor + factor(salmon01) +
               s(ydayCumul, bs = 'cr'), 
               data = dataIn, 
               method = "REML")

@@

@@

@@

predGrid <- expand.grid(riverFactor = unique(dataIn$riverFactor), 
                           cohortFactor = unique(dataIn$cohort),
                           ydayCumul = seq(250,1000, by = 10),
                           salmon01 = c(0,1),
                           cohortStrengthScaled = 0, #seq(-1.5,1.5, by = 0.5),
                           meanTemperatureScaledBySeason = 0, #seq(-1.5,1.5, by = 0.5),
                           meanFlowScaledBySeason = 0) %>%
                    left_join(., yDayAll) %>%
                    mutate(cohortRiverFactor = paste(cohortFactor, riverFactor, sep = '_'),
                           seasonRiverFactor = paste(season, riverFactor, sep = '_'),
                           cohort = cohortFactor,
                           #river = riverFactor,
                           meanTemperatureFlowScaledBySeason = meanTemperatureScaledBySeason * meanFlowScaledBySeason)

# predGridBKT2 <- expand.grid(riverFactor = unique(dataIn$riverFactor), 
#                            
#                            ydayCumul = seq(250,1000, by = 10),
#                            salmon01 = c(0,1)) %>%
#                     left_join(., yDayAll) %>%
#                     mutate()

pred <- predict(mod6_YDCcr_Cr_CSr_Tr_Fr_TFr, newdata = predGrid)#, type = 'lpmatrix')
#predBKT2 <- predict(mod1, newdata = predGridBKT2)#, type = 'lpmatrix')

predGrid2 <- bind_cols(predGrid, as.data.frame(pred)) %>% mutate(pred = pred)
#predGrid2BKT2 <- bind_cols(predGridBKT2, as.data.frame(predBKT2)) %>% mutate(pred = predBKT2, species = 'bkt')

@@

#using salmon01 * river because we need diff effects per river - even though AIC is worse for this model
modBKT <- bam(observedLength ~ riverFactor + factor(salmon01) +              
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
               s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor) +
               s(cohort, bs = 'cr', k = kIn, by = riverFactor) + 
               s(cohortStrengthScaled, bs = 'cr', k = kIn, by = riverFactor) +  # this is conspecific cS only
               s(meanTemperatureScaledBySeason, bs = 'cr', k = kIn, by = riverFactor) +
               s(meanFlowScaledBySeason, bs = 'cr', k = kIn, by = riverFactor) +
               s(meanTemperatureFlowScaledBySeason, bs = 'cr', k = kIn, by = riverFactor), 
             data = vB %>% filter(cohort %in% 2000:2013, species %in% 'bkt'), method = "REML")

@@

predGridBKT <- expand.grid(riverFactor = unique(modBKT$model$riverFactor), 
                           cohortFactor = unique(modBKT$model$cohort),
                           ydayCumul = seq(250,1000, by = 10),
                           salmon01 = c(0,1),
                           cohortStrengthScaled = 0, #seq(-1.5,1.5, by = 0.5),
                           meanTemperatureScaledBySeason = 0, #seq(-1.5,1.5, by = 0.5),
                           meanFlowScaledBySeason = 0) %>%
                    left_join(., yDayAll) %>%
                    mutate(cohortRiverFactor = paste(cohortFactor, riverFactor, sep = '_'),
                           seasonRiverFactor = paste(season, riverFactor, sep = '_'),
                           cohort = cohortFactor,
                           #river = riverFactor,
                           meanTemperatureFlowScaledBySeason = meanTemperatureScaledBySeason * meanFlowScaledBySeason)

@@

modBNT <- bam(observedLength ~ riverFactor + factor(salmon01) +              
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
               s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor) +
               s(cohort, bs = 'cr', k = kIn, by = riverFactor) + 
               s(cohortStrengthScaled, bs = 'cr', k = kIn, by = riverFactor) +  # this is conspecific cS only
               s(meanTemperatureScaledBySeason, bs = 'cr', k = kIn, by = riverFactor) +
               s(meanFlowScaledBySeason, bs = 'cr', k = kIn, by = riverFactor) +
               s(meanTemperatureFlowScaledBySeason, bs = 'cr', k = kIn, by = riverFactor), 
             data = vB %>% filter(cohort %in% 2000:2013, species %in% 'bnt'), method = "REML")

@@

predGridBNT <- expand.grid(riverFactor = unique(modBNT$model$riverFactor), 
                           cohortFactor = unique(modBNT$model$cohort),
                           ydayCumul = seq(250,1000, by = 10),
                           salmon01 = c(0,1),
                           cohortStrengthScaled = 0, #seq(-1.5,1.5, by = 0.5),
                           meanTemperatureScaledBySeason = 0, #seq(-1.5,1.5, by = 0.5),
                           meanFlowScaledBySeason = 0) %>%
                    left_join(., yDayAll) %>%
                    mutate(cohortRiverFactor = paste(cohortFactor, riverFactor, sep = '_'),
                           seasonRiverFactor = paste(season, riverFactor, sep = '_'),
                           cohort = cohortFactor,
                           #river = riverFactor,
                           meanTemperatureFlowScaledBySeason = meanTemperatureScaledBySeason * meanFlowScaledBySeason)

>>>>

mod6 <- bam(observedLength ~ riverFactor + factor(salmon01) +
               s(ydayCumul, bs = 'ps'), 
               data = dataIn, 
               method = "REML")

@@

# need to run code at top of 'predictions' block before running this chunk
predGrid2 <- expand.grid(riverFactor = unique(dataIn$riverFactor),
                           ydayCumul = seq(250,1000, by = 10),
                           salmon01 = c(0,1),
                         season = 2) %>%
                    #left_join(., yDayAll) %>%
                    mutate()

pred2 <- predict(mod6, newdata = predGrid2)#, type = 'lpmatrix')

predGrid2 <- bind_cols(predGrid2, as.data.frame(pred2)) %>% mutate(pred = pred2, species = 'bkt')

       aes(ydayCumul, pred)) +
  #geom_point() + 
  geom_vline(xintercept = c(365,365*2), color = 'grey60', linetype = 2) +
  geom_line(size = 1.1) +
  labs(y = 'Predicted body size (mm)',
       x = 'Days since age-0 Jan 1',
       color = 'Cohort') +
  scale_color_viridis_c(breaks = c(2000,2004,2008,2012), labels = c(2000,2004,2008,2012)) +
  theme_publication() +
  theme(legend.position = "right", 
        legend.direction = "vertical",
        legend.title = element_text(face = "plain")) +
  guides(color = guide_colorbar(barheight = 10,
                                barwidth = 1.5,
                                ticks = FALSE,
                                nbin = 100)) +
#  guide_legend(title = 'Cohort') +
  facet_wrap(~riverFactor)#, scales = 'free_y')

@@

```{r models without T and F}
# Not sure how to get models with season-specific T and F effects to work within season without messing up predictions. # Leave T and F out and focus on densities

mod6_YDCcr_Cr_CSr <- bam(observedLength ~ riverFactor + factor(salmon01) +              
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
               s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor) +
               s(cohort, bs = 'cr', k = kIn, by = riverFactor) + 
               s(cohortStrengthScaled, bs = 'cr', k = kIn, by = riverFactor),  # this is conspecific cS only
               data = dataIn, 
               method = "REML")

system.time(
mod6_YDCcr_Cr_CS3r <- bam(observedLength ~ riverFactor + factor(salmon01) +              
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
               s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor) +
               s(cohort, bs = 'cr', k = kIn, by = riverFactor) + 
               s(cohortStrengthScaledBKT, bs = 'cr', k = kIn, by = riverFactor) +
               s(cohortStrengthScaledBNT, bs = 'cr', k = kIn, by = riverFactor) +
               s(cohortStrengthScaledATS, bs = 'cr', k = kIn, by = riverFactor),  
               data = dataIn, 
               method = "REML")
)

system.time(
mod6_YDCcr_Cr <- bam(observedLength ~ riverFactor + factor(salmon01) +              
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
               s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor) +
               s(cohort, bs = 'cr', k = kIn, by = riverFactor),  
               data = dataIn, 
               method = "REML")
)

system.time(
mod6_YDCcr <- bam(observedLength ~ riverFactor + factor(salmon01) +              
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
               s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor),  
               data = dataIn, 
               method = "REML")
)

system.time(
mod6_YDC <- bam(observedLength ~ riverFactor + factor(salmon01) +              
               s(ydayCumul, bs = 'cr'),  
               data = dataIn, 
               method = "REML")
)

system.time(
mod6_YDCcr_Cr_CSBKTr <- bam(observedLength ~ riverFactor + factor(salmon01) +              
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
               s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor) +
               s(cohort, bs = 'cr', k = kIn, by = riverFactor) + 
               s(cohortStrengthScaledBKT, bs = 'cr', k = kIn, by = riverFactor),  
               data = dataIn, 
               method = "REML")
)

system.time(
mod6_YDCcr_Cr_CSBNTr <- bam(observedLength ~ riverFactor + factor(salmon01) +              
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
               s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor) +
               s(cohort, bs = 'cr', k = kIn, by = riverFactor) + 
               s(cohortStrengthScaledBNT, bs = 'cr', k = kIn, by = riverFactor),  
               data = dataIn, 
               method = "REML")
)

system.time(
mod6_YDCcr_Cr_CSATSr <- bam(observedLength ~ riverFactor + factor(salmon01) +              
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
               s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor) +
               s(cohort, bs = 'cr', k = kIn, by = riverFactor) + 
               s(cohortStrengthScaledATS, bs = 'cr', k = kIn, by = riverFactor),  
               data = dataIn, 
               method = "REML")
)

AIC(mod6_YDCcr_Cr_CSr, mod6_YDCcr_Cr_CS3r, mod6_YDCcr_Cr_CSBKTr, mod6_YDCcr_Cr_CSBNTr, mod6_YDCcr_Cr_CSATSr, mod6_YDCcr_Cr,mod6_YDCcr,mod6_YDC)  %>% rownames_to_column() %>% arrange(AIC)

summary(mod6_YDCcr_Cr_CSr)
plot(mod6_YDCcr_Cr_CS3r, pages = 7, shade = TRUE, scale = 0)

# bkt
#                rowname        df      AIC
# 1    mod6_YDCcr_Cr_CSr 173.85954 413790.2
# 2   mod6_YDCcr_Cr_CS3r 188.68933 415278.2
# 3 mod6_YDCcr_Cr_CSBKTr 167.72761 416582.2
# 4 mod6_YDCcr_Cr_CSBNTr 168.43917 417098.5
# 5 mod6_YDCcr_Cr_CSATSr 167.84300 417569.6
# 6        mod6_YDCcr_Cr 160.60559 417764.7
# 7           mod6_YDCcr 149.80639 420935.0
# 8             mod6_YDC  14.99337 429769.2

# bnt
#                rowname       df      AIC
# 1   mod6_YDCcr_Cr_CS3r 96.91663 231844.1
# 2    mod6_YDCcr_Cr_CSr 84.99160 231875.4
# 3 mod6_YDCcr_Cr_CSBKTr 84.39999 232019.6
# 4 mod6_YDCcr_Cr_CSBNTr 87.25717 232120.2
# 5 mod6_YDCcr_Cr_CSATSr 85.68893 232203.3
# 6        mod6_YDCcr_Cr 81.27554 232235.7
# 7           mod6_YDCcr 75.00492 233885.5
# 8             mod6_YDC 12.99685 238636.2
```

@@

predGrid <- expand.grid(riverFactor = unique(dataIn$riverFactor), 
                        cohortFactor = unique(dataIn$cohort),
                        ydayCumul = seq(250,1000, by = 10),
                        salmon01 = c(0,1),
                        cohortStrengthScaled = seq(-2.5,2.5, by = 0.25)) %>%
                    #left_join(., yDayAll) %>%
                    mutate(cohortRiverFactor = paste(cohortFactor, riverFactor, sep = '_'),
                           cohort = cohortFactor)

pred <- predict(mod6_YDCcr_Cr_CSr, newdata = predGrid)#, type = 'lpmatrix')

predGrid2 <- bind_cols(predGrid, as.data.frame(pred)) %>% mutate(pred = pred)

ggplot(predGrid2 %>% filter(salmon01 == 0, cohortStrengthScaled == 0),

@@

#using salmon01 * river because we need diff effects per river - even though AIC is worse for this model
modBKT <- bam(observedLength ~ riverFactor + factor(salmon01) +              
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
               s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor) +
               s(cohort, bs = 'cr', k = kIn, by = riverFactor) + 
               s(cohortStrengthScaled, bs = 'cr', k = kIn, by = riverFactor), 
             data = vB %>% filter(cohort %in% 2000:2013, species %in% 'bkt'), method = "REML")

@@

predGridBKT <- expand.grid(riverFactor = unique(dataIn$riverFactor), 
                        cohortFactor = unique(dataIn$cohort),
                        ydayCumul = seq(250,1000, by = 30),
                        salmon01 = c(0,1),
                        cohortStrengthScaled = seq(-2.5,2.5, by = 0.25)) %>%
                    #left_join(., yDayAll) %>%
                    mutate(cohortRiverFactor = paste(cohortFactor, riverFactor, sep = '_'),
                           cohort = cohortFactor)

@@

modBNT <- bam(observedLength ~ riverFactor + factor(salmon01) +              
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
               s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor) +
               s(cohort, bs = 'cr', k = kIn, by = riverFactor) + 
               s(cohortStrengthScaledBKT, bs = 'cr', k = kIn, by = riverFactor) +
               s(cohortStrengthScaledBNT, bs = 'cr', k = kIn, by = riverFactor) +
               s(cohortStrengthScaledATS, bs = 'cr', k = kIn, by = riverFactor), 
             data = vB %>% filter(cohort %in% 2000:2013, species %in% 'bnt'), method = "REML")

@@

predGridBNT <- expand.grid(riverFactor = unique(dataIn$riverFactor), 
                           cohortFactor = unique(dataIn$cohort),
                           ydayCumul = seq(250,1000, by = 30),
                           salmon01 = c(0,1),
                           cohortStrengthScaledBKT = seq(-2.5,2.5, by = 0.5),
                           cohortStrengthScaledBNT = seq(-2.5,2.5, by = 0.5),
                           cohortStrengthScaledATS = seq(-2.5,2.5, by = 0.5)) %>%
                    #left_join(., yDayAll) %>%
                    mutate(cohortRiverFactor = paste(cohortFactor, riverFactor, sep = '_'),
                           cohort = cohortFactor)
